PAMELA: an annotation-based Java Modeling Framework

### 2021-04-21 ResponseLetterMinor.tex

% Title:    A LaTeX Template For Responses To a Referees' Reports
% Author:   Petr Zemek <s3rvac@gmail.com>
% Homepage: https://blog.petrzemek.net/2016/07/17/latex-template-for-responses-to-referees-reports/
% License:  CC BY 4.0 (https://creativecommons.org/licenses/by/4.0/)
\documentclass[10pt]{article}

@@

\bigskip
Sincerely yours,

\bigskip
Sylvain, Guillaume, Caine, Joel, Jean-Christophe, Salvador, Fabien and Antoine

\pagebreak

\Reviewer{\#1}
\begin{response}{There remain or are introduced however, a substantial number of typos, language errors, and other minor issues that need correcting:}

All highlighted typos, languages errors and minor issues have been fixed.

\begin{itemize}

\item "to /avoid" on p3 (DONE)
\item "detailed thereafter" $->$ "detailed hereafter" (It's Correct : As adverbs the difference between hereafter and thereafter is that hereafter is in time to come; in some future time or state while thereafter is after that, from then on.)
\item "increasing PAMELA metamodel" $->$ "by \_extending\_ PAMELA\_'s\_ metamodel" (DONE)
\item "with setter/getter": make both plural (DONE)
\item POJO used here and elsewhere yet the acronym is introduced much later (DONE)
\item p4 l11-12: reduce etc. should be reduce\_s\_ etc. (Keep as it is)
\item odd spacing before : consistently throughout bullet list + text  (Keep as it is: trust Latex)
\item "in the section 3.2" $->$ "in Section 3.2" (style: named references deserve a capital and no article before) (DONE)
\item p5 and many times: "PAMELA metamodel", "PAMELA model", "PAMELA framework" need "the" before them. (Done ; with Pamela $->$ PAMELA)
\item "defines API": insert "an" (DONE)
\item "in Java interface": insert "a" (DONE)
\item "returning desired value", "taking value" etc.: insert "the" (DONE)
\item "May also declares"... (DONE)
\item p6 l46: @Embedded/"as embedded" meaning?
\item p7 "of launched Java application" missing "the" (DONE)
\item "is/are" $->$ "is (are)", in line with parentheses earlier in sentence (DONE)
\item "Experimented programmers": I doubt it, unless you got major ethical clearance (Replaced by "Experienced PAMELA programmers")
\item "Listing 2: model" $->$ "~: Model" consistent with Listing 1 (DONE)
\item the book title is "Lord of the ring\_s\_" with s (DONE)
\item "using classical Java ... scheme": missing "the" (DONE)
\item "Java programming language": insert "The" (DONE)
\item l43 inconsistent spacing (DONE)
\item l46 package ... package (DONE)
\item p11 l36 sticking into margin (Final check to do)
\item S.4.3 please add size in LoC and classes (?)
\item java $->$ Java (DONE)
\item p13 "diana component": insert "the" (DONE)
\item "experimentation" $->$ "experiment" (DONE)
\item "When Pamela has been" $->$ "After Pamela was" (DONE : if we consider it's over)
\item l36 the quotation mark before code generation is the wrong one 'changed to ' and ' ?)
\item p 14 "appertaining" $->$ "pertaining", "relating" (Kept as is. As verbs the difference between pertain and appertain is that pertain is to belong while appertain is to belong to or be a part of, whether by right, nature, appointment, or custom; to relate to.)
\end{itemize}

\end{response}

@@

\pagebreak

\Reviewer{\#2}
\begin{response}{Would it be possible to structure it more, because currently it is several paragraphs the one bellow the others.}

Section 2 has been decomposed into 2 subsections

\end{response}

\begin{response}{figure 2 is not enough described in section 2 and even in the paper.}

The end of 2.1 subsection provides an explanation of figure 2.

\end{response}

\begin{response}{More over, concerning section 4.4, I appreciate this new section. However, while reading, it is not a good advertisement for Pamela. Overhead (x20) is huge. Perhaps it is not the right benchmark if it provides bad results that are not representing the right usage.}

SM comments

\end{response}

### 2021-04-21 pamelaProcess.tex
Adding minor revision response letter
restructuration section 2

We advocate for a strong coupling between model and sources code, to give architects and developers a way to both interact during the whole development cycle. PAMELA is an annotation-based Java modeling framework providing a smooth integration between model and code, without code generation nor externalized model serialization. Instead of generating the code, the API (mostly Java interfaces with abstract method declarations) is locally executed (interpreted). The idea is to avoid separation between modeling and code to facilitate consistency management and to avoid round-trip issues. Figure \ref{fig:PamelaVision} summarizes the PAMELA architecture and is more detailed thereafter.

@@

%\subsection{PAMELA Modeling/Development Process}

Coupling model and code into the same artifact opens new ways of programming. The classical (metadata enabled) programming process relies on \emph{programmers} that produce code reusing pre-existing modeling concepts. These concepts are implemented by \emph{modelers} that provide the right annotations the programmers use. This is, for instance, the process followed by Jakarta EE (JEE) developers reusing JEE specific annotations. The evolution rhythm between models and code is low. This programming way is still possible with PAMELA, but we allow the ability to reach a high evolution rhythm when the programmer also becomes the modeler. In fact, when the programmer identifies a pattern, an abstraction, a generalization, s/he can use PAMELA to develop and capitalize on this abstraction by extending PAMELA's metamodel.

% Je pense qu'ici il faudrait plutot parler du modèle. Le métamodèle PAMELA est celui représenté figure 3. Ou alors parler du metametamodele PAMELA

The developed metamodels are implemented by annotations that rely on Java/JVM entities and mechanisms. They include consistency checking, which constrain their use and helps the programmer to avoid inconsistencies or errors. We have first experimented their use with setters/getters to define Plain Old Java Objects (POJO), with traits to implement multiple inheritance or with roles and rules to set security rules on classes.

>>>>

We advocate for a strong coupling between model and sources code, to give architects and developers a way to both interact during the whole development cycle.

\subsection{Architecture overview}

PAMELA is an annotation-based Java modeling framework providing a smooth integration between model and code, without code generation nor externalized model serialization. Instead of generating the code, the API (mostly Java interfaces with abstract method declarations) is locally executed (interpreted). The idea is to avoid separation between modeling and code to facilitate consistency management and to avoid round-trip issues. Figure \ref{fig:PamelaVision} summarizes the PAMELA architecture.
The left side shows the structure of the application source code where java files contain the code and annotations that link part of the code to PAMELA model entities such as class, attribute, method, etc. The right side shows that at runtime the PAMELA interpreter maintains the relationships between the application binary, result of the java compilation, and the internal representation of PAMELA models. The preservation of the structure and of the whole information allow a very good and high level of control over the execution. In the following section, we present some indications on the different ways of programming using PAMELA.

@@

%\subsection{PAMELA Modeling/Development Process}
\subsection{Usage of PAMELA}

Coupling model and code into the same artifact opens new ways of programming. The classical (metadata enabled) programming process relies on \emph{programmers} that produce code reusing pre-existing modeling concepts. These concepts are implemented by \emph{modelers} that provide the right annotations the programmers use. This is, for instance, the process followed by Jakarta EE (JEE) developers reusing JEE specific annotations. 
% Joel : une ref pour Jakarta ?
The evolution rhythm between models and code is low. This programming way is still possible with PAMELA, but we allow the ability to reach a high evolution rhythm when the programmer also becomes the modeler. In fact, when the programmer identifies a pattern, an abstraction, a generalization, s/he can use PAMELA to develop and capitalize on this abstraction by extending PAMELA's metamodel. 
% Joel : le MM de Pamela n'est pas étendu en tant que tel (=M3) c'est le niveau M2 qui étendu ??? mais la bibliothèque de metamodèles est étendue ??? donc juste mettre PAMELA's metamodels. ????
% Je pense qu'ici il faudrait plutot parler du modèle. Le métamodèle PAMELA est celui représenté figure 3. Ou alors parler du metametamodele PAMELA

The developed metamodels are implemented by annotations that rely on Java/JVM entities and mechanisms. They include consistency checking, which constrain their use and help the programmer to avoid inconsistencies or errors. We have first experimented their use with setters/getters to define Plain Old Java Objects (POJO), with traits to implement multiple inheritance or with roles and rules to set security rules on classes.
